Symbolic Abstract Data Type Inference

### 2015-07-07 histories.tex
example <= relation
Fixed space hacks.
Improved histories section.

  Consider the following three executions of a single-value register object
  with read and write methods
  \begin{center}
    \setlength{\tabcolsep}{1em}
    \begin{tabular}{l|l|l}
      $e_1$                           & $e_2$                           & $e_3$ \\
      $1:\mathrm{write}(a)$           & $1:\mathrm{write}(a)$           & $1:\mathrm{write}(a)$ \\
      $1:\mathbf{ret}$                & $2:\mathrm{write}(b)$           & $1:\mathbf{ret}$ \\
      $2:\mathrm{write}(b)$           & $1:\mathbf{ret}$                & $2:\mathrm{write}(b)$ \\
      $2:\mathbf{ret}$                & $2:\mathbf{ret}$                & $3:\mathrm{read}$ \\
      $3:\mathrm{read}$               & $3:\mathrm{read}$               & $2:\mathbf{ret}$ \\
      $3:\mathbf{ret} \Rightarrow a$  & $3:\mathbf{ret} \Rightarrow a$  & $3:\mathbf{ret} \Rightarrow a$ \\
    \end{tabular}
  \end{center}
  in which actions are preceded by their operation identifiers. Executions
  $e_2$ and $e_3$ are obtained from $e_1$ by permuting the return actions of
  operations $1$ and $2$, respectively, forward, or symmetrically, the call
  actions of operations $2$ and $3$, respectively, backward. Thus while the
  operations of Execution $e_1$ are sequential, each following the previous in
  time, those of $e_2$ and $e_3$ overlap. While $e_1$ should not be admitted
  by an \emph{atomic} register, since the read of Operation~$3$ does not
  return the most-recently-written value, both $e_2$ and $e_3$ could be
  admitted, since there exist linearizations of the partially-ordered operations
  such that Operation~$3$ does return the most-recently-written value.

@@

In this work, histories are abstractions of executions which retain method
names yet lose the exact argument and return values, and retain the relative
order of operations yet lose the exact sequence of call and return actions.
Formally, a \emph{history} is a tuple $h = \tup{O,<,c,f,m,r}$ where
\begin{itemize}

@@

>>>>

  Consider the following three executions of a single-value register object
  with read and write methods:
  \begin{center}
    \hfill
    \begin{minipage}{0.3\linewidth}
      $e_1$
      \begin{verbatim}
1: call write(a)
1: return
2: call write(b)
2: return
3: call read
3: return => a
      \end{verbatim}
    \end{minipage}
    \hfill
    \begin{minipage}{0.3\linewidth}
      $e_2$
      \begin{verbatim}
1: call write(a)
2: call write(b)
1: return
2: return
3: call read
3: return => a
      \end{verbatim}
    \end{minipage}
    \hfill
    \begin{minipage}{0.3\linewidth}
      $e_3$
      \begin{verbatim}
1: call write(a)
1: return
2: call write(b)
3: call read
2: return
3: return => a
      \end{verbatim}
    \end{minipage}
    \hfill
  \end{center}
  Operation identifiers precede actions. Executions~$e_2$ and~$e_3$ are obtained
  from $e_1$, respectively, by permuting the return actions of Operations~$1$
  and~$2$ and the call actions of operations $2$ and $3$. Thus while the
  operations of Execution~$e_1$ are sequential, each following the previous in
  time, those of $e_2$ and $e_3$ overlap. While $e_1$ should not be admitted by
  an atomic register, since the read of Operation~$3$ does not return the
  most-recently-written value, both $e_2$ and $e_3$ should be admitted, since
  Operation~$3$ returns the most-recently-written value in some linearization of
  the overlapping operations.

@@

Histories abstract executions, retaining method
names yet losing exact argument and return values, and retaining the relative
order of operations, yet losing the exact sequence of call and return actions.
Formally, a \emph{history} is a tuple $h = \tup{O,<,c,f,m,r}$ where
\begin{itemize}

@@

\begin{example}

  By unordering the first two operations of the history
\begin{verbatim}
    [1:1] write(a)        #
    [2:2] write(b)          #
    [3:2] read => b (RO)      #
\end{verbatim}
  we derive the $\to_\mathrm{o}$-related history
\begin{verbatim}
    [1:1] write(a)        #
    [2:2] write(b)        #
    [3:2] read => b (RO)    #
\end{verbatim}
  from which we can derive the $\to_\mathrm{c}$-related history
\begin{verbatim}
    [1:1] write(a)        #
    [2:2] write(b)        #
    [3:_] read* (RO)        #
\end{verbatim}
  by making Operation~$3$ pending, and from which we can derive the
  $\to_\mathrm{p}$-related history
\begin{verbatim}
    [1:1] write(a)        #
    [2:2] write(b)        #
    [3:_] read* (RO)        #
    [4:4] write(c)*         #
\end{verbatim}
  by adding an additional pending operation.

Since implementations only dictate methods' executions between their respective
calls and returns, for any execution they admit, they must also admit executions
with weaker inter-operation ordering, in which calls may happen earlier, and/or
returns later. Thus any weakening of an admitted history must also be admitted.
